---
title: "DAC Analytics: Global Direct Air Capture Facilities"
author: "Nobel Girmay"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    code_folding: hide
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)

library(tidyverse)
library(plotly)
library(knitr)
library(kableExtra)
library(leaflet)
library(lubridate)
```

## Overview

This report analyzes a curated database of **real-world Direct Air Capture (DAC)
facilities** worldwide, spanning operational plants, projects under construction,
and announced installations. Data is sourced from the IEA, the U.S. DOE DAC Hubs
program, and company disclosures.

The companion interactive dashboard (built in R Shiny with Leaflet and Plotly)
is in the [project repository](https://github.com/ngirmay/Carbon_footprints).

```{r load-data}
# Load the real facilities database
dac <- read_csv("data/dac_facilities.csv", show_col_types = FALSE) %>%
  mutate(
    start_date = as.Date(start_date),
    status = factor(status, levels = c("Operational", "Under Construction", "Planned"))
  )

n_fac      <- nrow(dac)
n_op       <- sum(dac$status == "Operational", na.rm = TRUE)
total_cap  <- sum(dac$capacity_tonnes_yr, na.rm = TRUE)
op_cap     <- sum(dac$capacity_tonnes_yr[dac$status == "Operational"], na.rm = TRUE)
```

**`r n_fac`** facilities tracked across **`r n_distinct(dac$country)`** countries.
**`r n_op`** are operational today, with a combined pipeline capacity of
**`r format(total_cap, big.mark = ",")`** tonnes CO&#8322;/year
(**`r format(op_cap, big.mark = ",")`** tonnes/year already operational).

## Global Map

```{r map}
pal <- colorFactor(c("#2f5e28", "#3F7D34", "#8BC34A"), domain = dac$status)

leaflet(dac) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    ~longitude, ~latitude,
    radius = ~pmax(4, log10(capacity_tonnes_yr) * 2.5),
    color = ~pal(status), fillOpacity = 0.75, stroke = FALSE,
    popup = ~paste0(
      "<b>", name, "</b><br>", company, "<br>",
      format(capacity_tonnes_yr, big.mark = ","), " t/yr &middot; ", status, "<br>",
      technology
    )
  ) %>%
  addLegend("bottomright", pal = pal, values = ~status, title = "Status")
```

## Capacity by Facility

```{r capacity-bar}
dac %>%
  arrange(desc(capacity_tonnes_yr)) %>%
  slice_head(n = 15) %>%
  plot_ly(
    x = ~capacity_tonnes_yr,
    y = ~reorder(name, capacity_tonnes_yr),
    color = ~status,
    colors = c("#2f5e28", "#3F7D34", "#8BC34A"),
    type = "bar", orientation = "h"
  ) %>%
  layout(
    title = "Top 15 DAC Facilities by Announced Capacity",
    xaxis = list(title = "Capacity (tonnes CO2 / year)"),
    yaxis = list(title = ""),
    margin = list(l = 160)
  )
```

## Pipeline by Status and Region

```{r status-region}
dac %>%
  group_by(region, status) %>%
  summarise(capacity = sum(capacity_tonnes_yr), .groups = "drop") %>%
  plot_ly(
    x = ~region, y = ~capacity, color = ~status,
    colors = c("#2f5e28", "#3F7D34", "#8BC34A"), type = "bar"
  ) %>%
  layout(
    barmode = "stack",
    title = "Announced Capacity by Region and Status",
    xaxis = list(title = ""),
    yaxis = list(title = "Capacity (tonnes CO2 / year)")
  )
```

## Technology Mix

```{r tech-table}
dac %>%
  count(technology, wt = capacity_tonnes_yr, name = "capacity") %>%
  arrange(desc(capacity)) %>%
  mutate(capacity = paste0(format(capacity, big.mark = ","), " t/yr")) %>%
  kable(col.names = c("Technology", "Total Announced Capacity")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Deployment Timeline

```{r timeline}
dac %>%
  filter(!is.na(start_date)) %>%
  arrange(start_date) %>%
  mutate(cumulative = cumsum(capacity_tonnes_yr)) %>%
  plot_ly(x = ~start_date, y = ~cumulative, type = "scatter", mode = "lines+markers",
          line = list(color = "#3F7D34"), marker = list(color = "#2f5e28")) %>%
  layout(
    title = "Cumulative Announced Capacity Over Time",
    xaxis = list(title = ""),
    yaxis = list(title = "Cumulative capacity (tonnes CO2 / year)")
  )
```

## Notes

- Capacities are **announced/nameplate** figures; operational output varies with
  load factor, energy source, and ambient conditions.
- "Planned" facilities are subject to final investment decisions and may shift.
- Data compiled from IEA, DOE DAC Hubs, and 1PointFive / Climeworks / Heirloom
  disclosures. See `data/dac_facilities.csv` for the full source column.
